---
title: "A&S 500 Final Project"
author:
  - name: "Jacob Saindon" #splitting should work in new versions of Quarto
    url: "https://jacographer.com/"
    orcid: 0000-0002-3808-9704
    affiliations:
      - name: "University of Kentucky"
citation:
  url: "https://jacographer.com/"
knitr:
  opts_chunk: 
    message: false
    warning: false
    fig-height: 5
    fig-width: 3
    fig-align: center
    cap-location: margin
execute:
  freeze: auto  # re-render only when source changes
format:
  html:
    theme: 
      - flatly
      - custom.scss
    toc: true
    toc-depth: 3
    toc-location: left #'body' goes into the top of doc
    toc-title: "Sections"
    title-block-banner: true
    #title-block-style: plain
    backgroundcolor: "green"
    fontcolor: "purple"
    include-in-header: google-font.html #file included in final proj.
    mainfont: Atkinson Hyperlegible
  docx: default
  pdf: default
    #toc: true
pdf-engine: pdflatex #necessary for this R server.
---

## Loading Packages

```{r}
#| eval: false
install.packages("gt")
install.packages("htmltools")
```

```{r}
library(tidyverse)
library(lubridate)
library(gt)
```

## Data

```{r}
snap <- read_csv("data/snap_political_20.csv")
```

## Tidying

```{r}
snap |>
  filter(
    CountryCode == "united states",
    !is.na(Interests),
    !is.na(`Regions (Included)`)
  ) |>
  mutate(
    STATE_INC = str_split(`Regions (Included)`, ",")
    ) |>
  unnest(cols = STATE_INC) -> snap_states

snap_states |>
  mutate(
    S_MONTH = month(StartDate, label = TRUE, abbr = TRUE)) -> snap_sm

snap_sm |>
  mutate(
    TARGET = str_split(`Interests`, ",")
    ) |>
    unnest(cols = TARGET) -> snap_unnest
```

```{r}
snap_unnest |>
  summarise(
      .by = c(STATE_INC,TARGET,S_MONTH),
      TOTAL = n()) |>
  mutate(
    id = row_number()) -> snap_tidy # all states, interests, and months in separate rows

snap_tidy
```

## Data Summaries

```{r}
snap_sm
```

```{r}
snap_sm |>
  group_by(STATE_INC) |>
  summarise(
    COUNT = n()) |>
  arrange(-COUNT) -> state_count

 # count of ads run by state
```

```{r}
snap_sm |>
    filter(
    STATE_INC == "North Carolina" | STATE_INC == "Arizona" | STATE_INC == "Georgia" | STATE_INC == "Florida" | STATE_INC == "Michigan" | STATE_INC == "Pennsylvania" | STATE_INC == "Wisconsin" | STATE_INC == "Virginia" | STATE_INC == "Maine" | STATE_INC == "Iowa") |>
  group_by(STATE_INC, S_MONTH)|>
  summarise(
    COUNT = n()) |>
  arrange(-COUNT) -> sm_count

sm_count # count of ads run by state, per month
```

```{r}
sm_count |>
ggplot(aes(x=S_MONTH, y=COUNT))+
  geom_col()+
  facet_wrap(~STATE_INC) # volume of ads launching per state
```

```{r}
snap_tidy |>
  group_by(STATE_INC, S_MONTH) |>
  summarise(TOTAL = sum(TOTAL)) |>
  arrange(-TOTAL) -> sm_sum # total number of audiences targeted by all ads, per state, per month

sm_sum
```

```{r}
snap_tidy |>
  group_by(TARGET, S_MONTH) |>
  summarise(
    SUM = sum(TOTAL)) |>
  arrange(-SUM) -> tm_sum # total number of ads including a given parameter, per month

tm_sum
```

```{r}
snap_tidy |>
  group_by(TARGET, STATE_INC) |>
  summarise(
    SUM = sum(TOTAL)) |>
  arrange(-SUM) -> ts_sum # total number of ads including a given parameter, per state

ts_sum
```

## Top Interests by State and Month

```{r}
snap_tidy |>
  slice_max(TOTAL, by = c(STATE_INC, S_MONTH)) -> max_list # all top interests per state, per month

max_list
```

```{r}
max_list |>
  group_by(STATE_INC, S_MONTH) |>
  subset(select = c(STATE_INC, S_MONTH, TARGET)) |>
    nest(.key = "TOP_TARGET") |>
  arrange(S_MONTH, STATE_INC) -> top_snap

top_snap # column with list (tibble) of top interests by per state, per month
```

## Plotting

```{r}
# month_max |> 
#   ggplot(aes(S_MONTH, MAX, fill = STATE_INC))+
#   geom_col()+
#   theme(legend.position="none") 
```

```{r}
#data |>
#group by (month, state, interest) |>
#summarise (n == n()) |>
#filter (n==max(n)) |>
#select(-month)
#maxes |>
#leftjoin(data)
```

Text analysis of targeting categories, by state.

Unnest both "Regions" and "Interest" fields to get tidy.

#dlxs are oracle datalogix audiences? https://businesshelp.snapchat.com/s/article/custom-audiences?language=en_US

targetsmart and i360: https://www.axios.com/2022/09/08/snap-voter-data-republican-democrats

Geographic Scales:

1.  Region (State)
    -   Inclusions

    -   Exclusions

    -   Total Ads

    -   Total Spend

    -   Total Impressions

    -   Interests (text analysis)
2.  Metros (City)
    -   Inclusions

    -   Exclusions

    -   Total Ads

    -   Total Spend

    -   Total Impressions

    -   Interests (text analysis)
3.  Postal Codes (Zip)
    -   Inclusions

    -   Exclusions

    -   Total Ads

    -   Total Spend

    -   Total Impressions

    -   Interests (text analysis)
4.  Location Categories
    -   Inclusions

    -   Exclusions

    -   Total Ads

    -   Total Spend

    -   Total Impressions

    -   Interests (text analysis)
5.  Electoral Districts
    -   Inclusions

    -   Exclusions

    -   Total Ads

    -   Total Spend

    -   Total Impressions

    -   Interests (text analysis)
